Lesson 4
========================================================

```{r}
install.packages("ggplot2")


library(ggplot2)
data(diamonds) 
dim(diamonds) # 53940 obs and 10 variables
str(diamonds) # 3 Ordered Factors
str(diamonds$color)
?diamonds # D Represents the best Color

ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price Histogram")

summary(diamonds$price)

nrow(subset(diamonds, price < 500))
?nrow
ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 10, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price Histogram between $0 and $1500.") + 
  coord_cartesian(xlim=c(0,1500))

ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_grid(. ~ cut)
?diamonds
diamonds[which.max(diamonds$price),]
#or
summary(diamonds$price) # Max = 18820
diamonds[diamonds$price == 18823,] # 18820 doesn't exist!?!?!?

diamonds[which.min(diamonds$price),]

summary(diamonds$price) # Min = 326
diamonds[diamonds$price == 326,]
by(diamonds$price,diamonds$cut,summary)


ggplot(diamonds) + geom_histogram(aes(x=price), binwidth = 100, 
                                  color = "black", fill = "lightblue") + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_grid(. ~ cut)

ggplot(diamonds) + geom_histogram(aes(x=price)) + 
  ggtitle("Diamonds Price by Cut Histogram") + 
  facet_wrap(~ cut,scales="free_y",ncol = 5)

qplot(x = price, data = diamonds) + facet_wrap(~cut, scales="free")

ggplot(diamonds) + geom_histogram(aes(x=price/carat), binwidth = 0.05,
                                  color = "black", fill = "lightblue") +
  ggtitle("Histogram of Price per Carat, facet by Cut.") + 
  scale_x_log10() +
  facet_grid(. ~ cut)

by(diamonds$price, diamonds$color, summary)
ggplot(diamonds) + geom_boxplot(aes(x = color,  y = price, fill = color)) + 
  coord_cartesian(ylim = c(0, 8000)) + 
  scale_y_continuous(breaks = seq(0,8000,500)) +
  ggtitle("Diamonds Price by Color.")


by(diamonds$price, diamonds$color, summary)
by(diamonds$price, diamonds$color, IQR)

ggplot(diamonds) + geom_boxplot(aes(x=color,y=price/carat, fill=color)) + 
  scale_y_continuous(breaks = seq(0,8000,500)) + 
  coord_cartesian(ylim = c(0,8000)) + 
  ylab("Price per Carat") +
  ggtitle("Price per Carat by Color")

ggplot(diamonds) + geom_freqpoly(aes(x = carat), binwidth = 0.02) + 
  scale_x_continuous(breaks = seq(0,5,0.1)) + 
  geom_hline(y = 2000,color = "red") + 
  ggtitle("Carat Frequency Polygon")

install.packages("swirl")

library("swirl")
swirl()


```

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

install.packages("dplyr")
library(dplyr)

age_groupes <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groupes,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n()
)

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)

ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) +
  geom_line()
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
pf
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) +
  geom_point()
  
```

***

### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}

```

Create your plot!

```{r Conditional Means Plot}

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

