---
output:
  html_document: default
  pdf_document: default
---
Red Wine Quality by Hammad Al-Hammad
========================================================

> **What is the Red Wine?**: Red wine is a type of wine made from dark-colored (black) grape varieties. The actual color of the wine can range from intense violet, typical of young wines, through to brick red for mature wines and brown for older red wines. The juice from most purple grapes is greenish-white; the red color comes from anthocyan pigments (also called anthocyanins) present in the skin of the grape; exceptions are the relatively uncommon teinturier varieties, which produce a red colored. Much of the red-wine production process therefore involves extraction of color and flavor components from the grape skin.


> **What is the dataset?**: This tidy data set contains 1,599 red wines with 11 variables on the chemical properties of the wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).


> **Data Overview:** This dataset is public available for research. The details are described in [Cortez et al., 2009]. 

   Input variables (based on physicochemical tests):
   
   1 - fixed acidity (tartaric acid - g / dm^3)
   
   2 - volatile acidity (acetic acid - g / dm^3)
   
   3 - citric acid (g / dm^3)
   
   4 - residual sugar (g / dm^3)
   
   5 - chlorides (sodium chloride - g / dm^3
   
   6 - free sulfur dioxide (mg / dm^3)
   
   7 - total sulfur dioxide (mg / dm^3)
   
   8 - density (g / cm^3)
   
   9 - pH
   
   10 - sulphates (potassium sulphate - g / dm3)
   
   11 - alcohol (% by volume)
   
   12 - quality (score between 0 and 10)



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

# Load all libraries 
library('ggplot2')
library('knitr')
library('dplyr')
library(GGally)
library(scales)
library(memisc)
library(gridExtra)

```


#### Now, We are going to load data into the workspace.Then, we are going to print the internal structure of the dataset.


```{r echo=FALSE, Load_the_Data}

# Load the Data
rw <- read.csv('wineQualityReds.csv')
# Print structure of the dataset
str(rw)

```


As listed above, the data contains 1599 obs. of  13 variables. The first variable (X) is useles in our analysis. So, we are going to delete it.


```{r echo=FALSE, Remove_ID}

# Delete X from the dataset
rw$X <- NULL

```


#### Experts in alcohol always categorize them to 3 types of effectiveness:
1. Light: Alcohol < 9

2. Medium: Alcohol >= 9 and Alcohol <= 12 

3. Strong: Alcohol > 12

*So, We are going to add new column to categorize the wine type if it's Light, Medium or Strong.*


```{r echo=FALSE, Add_Category}

# Add Category to the dataset
rw$category <- ''
rw$category[rw$alcohol<=9] <- 'Light'
rw$category[9<rw$alcohol & rw$alcohol<=12] <- 'Medium'
rw$category[12<rw$alcohol] <- 'Strong'
rw$category <- as.factor(rw$category)

```


#### Let's see how many one of our sample in each category.

```{r echo=FALSE, Disiplay_Category_Table}

# Display a table which shows types of categories and their total
table(rw$category)

```


As we see, Light = 37, Medium = 1421 and Strong = 141


```{r echo=FALSE, Disiplay_QPlot_Category}

# Display a visualization which shows types of categories and their total
qplot(x=category, data=rw)

```


#### Experts in alcohol also ranking them to 3 types of quality:
1. Poor: Quality < 9

2. Good: Quality >= 9 and Quality <= 12 

3. Excellent: Quality > 12

*So, We are going to add new column to rank the wine type if it's Poor, Good or Excellent*


```{r echo=FALSE, Add_Ranking}

# Add Ranking to the dataset
rw$ranking[5>=rw$quality] <- 'Poor'
rw$ranking[5<rw$quality & rw$quality <8] <- 'Good'
rw$ranking[8<=rw$quality] <- 'Excellent'
rw$ranking <- as.factor(rw$ranking)

```


#### Let's see how many one of our sample in each ranking


```{r echo=FALSE, Disiplay_Ranking_Table}

# Display a table which shows types of ranking and their total
table(rw$ranking)

```


As we see, Excellent = 18, Good = 837 and Poor = 744


```{r echo=FALSE, Disiplay_QPlot_Ranking}

# Display a visualization which shows types of ranking and their total
qplot(x=ranking, data=rw)

```


# Univariate Plots Section


> **Tip**: Here, we are going to perform some preliminary exploration of Red Wines dataset. 


```{r echo=FALSE, Print_Summary_Of_RW}

# Print a summary about the data [Each Variable]
summary(rw)

```

The result above is listing basic statistics about each variable.


```{r echo=FALSE, Make_Plot_Function}

# This is a basic function, which takes a variable and histogram's bandwidth.
# Then return ggplot diagram.
make_plot <- function(variable, binwidth) {
  plot <- ggplot(aes_string(x = variable), data = rw) +
          geom_histogram(binwidth = binwidth)
  return(plot)
}

```


#### Here is a histogram about the fixed.acidity variable.

```{r echo=FALSE, Make_Plot_Function_For_fixed.acidity}

make_plot('fixed.acidity', 0.3) 

```

The fixed.acidity is skewed to the right. 


#### Here is a summary about the fixed.acidity variable.

```{r echo=FALSE, Disply_Summary_For_fixed.acidity}

summary(rw$fixed.acidity)

```

The above table shows a summary of the fixed.acidity variable.


#### Here is a histogram about the citric.acid variable.

```{r echo=FALSE, Make_Plot_Function_For_citric.acid}

make_plot('citric.acid', 0.05) 

```

The citric.acid also is skewed to the right. 


#### Here is a summary about the citric.acid variable.

```{r echo=FALSE, Disply_Summary_For_citric.acid}

summary(rw$citric.acid)

```

The above table shows a summary of the citric.acid variable.


#### Here is a histogram about the pH variable.

```{r echo=FALSE, Make_Plot_Function_For_pH}

make_plot('pH', 0.05) 

```

The pH is normally distributed. 


#### Here is a summary about the pH variable.

```{r echo=FALSE, Disply_Summary_For_pH}

summary(rw$pH)

```

The above table shows a summary of the pH variable.


#### Here is a histogram about the chlorides variable.

```{r echo=FALSE, Make_Plot_Function_For_chlorides}

make_plot('chlorides', 0.01) 

```

The chlorides is skewed to the right. *We are going to apply the 10th log to it.*

```{r echo=FALSE, Make_Plot_Function_For_chlorides_log10}

qplot(x=chlorides, data=rw)+
  scale_x_log10()


```

The distribution of chlorides is normally distributed.

#### Here is a summary about the chlorides variable.

```{r echo=FALSE, Disply_Summary_For_chlorides}

summary(rw$chlorides)

```

The above table shows a summary of the chlorides variable.

#### Here is a histogram about the residual.sugar variable.

```{r echo=FALSE, Make_Plot_Function_For_residual.sugar}

make_plot('residual.sugar', 0.05) 

```

The residual.sugar is skewed to the right. *We are going to apply the 10th log to it.*

```{r echo=FALSE, Make_Plot_Function_For_residual.sugar_log10}

qplot(x=residual.sugar, data=rw)+
  scale_x_log10()


```

The distribution of residual.sugar is normally distributed.


#### Here is a summary about the residual.sugar variable.

```{r echo=FALSE, Disply_Summary_For_residual_sugar}

summary(rw$residual.sugar)

```

The above table shows a summary of the residual.sugar variable.


#### Here is a histogram about the density variable.

```{r echo=FALSE, Make_Plot_Function_For_density}

make_plot('density', 0.001) 

```

The density is normally distributed. 


#### Here is a summary about the density variable.

```{r echo=FALSE, Disply_Summary_For_residual.sugar}

summary(rw$density)

```

The above table shows a summary of the density variable.

#### Here is a histogram and box plot about the alcohol variable.

```{r echo=FALSE, Make_Plot_For_alcohol}

# alchol plot
alchol_plot <- make_plot('alcohol', 1)
alchol_boxplot <- ggplot(aes(x=1, y=alcohol), data=rw) + geom_boxplot()

grid.arrange(alchol_plot, alchol_boxplot,nrow=1)

```

We can see that most of alcohol exists from 9 to 11.

#### Here is a summary about the alcohol variable.

```{r echo=FALSE, Disply_Summary_For_alcohol}

summary(rw$alcohol)

```

The above table shows a summary of the alcohol variable.


#### Here is a histogram about the free.sulfur.dioxide variable.

```{r echo=FALSE, Make_Plot_Function_For_free.sulfur.dioxide}

make_plot('free.sulfur.dioxide', 2) 

```

The free.sulfur.dioxide is skewed to the right. *We are going to apply the 10th log to it.*


```{r echo=FALSE, Make_Plot_Function_For_free.sulfur.dioxide_log10}

qplot(x=free.sulfur.dioxide, data=rw)+
  scale_x_log10()


```

The distribution of free.sulfur.dioxide isn't normally distributed. It seems to be like a bimodal distribution. Because free.sulfur.dioxide near to 9 is very low.


#### Here is a summary about the free.sulfur.dioxide variable.

```{r echo=FALSE, Disply_Summary_For_free.sulfur.dioxide}

summary(rw$free.sulfur.dioxide)

```

The above table shows a summary of the free.sulfur.dioxide variable.


#### Here is a histogram about the total.sulfur.dioxide variable.

```{r echo=FALSE, Make_Plot_Function_For_total.sulfur.dioxide}

make_plot('total.sulfur.dioxide', 2) 

```

The total.sulfur.dioxide is skewed to the right. *We are going to apply the 10th log to it.*


```{r echo=FALSE, Make_Plot_Function_For_total.sulfur.dioxide_log10}

qplot(x=total.sulfur.dioxide, data=rw)+
  scale_x_log10()


```

The distribution of total.sulfur.dioxide is normally distributed.

#### Here is a summary about the total.sulfur.dioxide variable.

```{r echo=FALSE, Disply_Summary_For_total.sulfur.dioxide}

summary(rw$total.sulfur.dioxide)

```

The above table shows a summary of the total.sulfur.dioxide variable.


#### Here is a histogram about the volatile.acidity variable.

```{r echo=FALSE, Make_Plot_Function_For_volatile.acidity}

make_plot('volatile.acidity', 0.1) 

```

The volatile.acidity is normally distributed.


#### Here is a summary about the volatile.acidity variable.

```{r echo=FALSE, Disply_Summary_For_volatile.acidity}

summary(rw$volatile.acidity)

```

The above table shows a summary of the volatile.acidity variable.





# Univariate Analysis


### What is the structure of your dataset?

Red Wine dataset contains 1599 records. Also, it has 14 variables. And this is the structure

```{r echo=FALSE, Disply_Names_Of_Variables}

names(rw)

```


### What is/are the main feature(s) of interest in your dataset?

For sure, main interest is in quality variable. We are going to see how it get affected by variables like percentage of alcohol, density and chlorides.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Category and ranking will help me in the investigation into my features of interest.

### Did you create any new variables from existing variables in the dataset?

Yes, as listed above. I created the *category* of wine [Light, Medium, Strong] and *rating* of wine [Poor, Good, Excellent]

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

Some variabels like [residual.sugar & alcohol] are skewed to the right. So, I made a log10 transformation to them to be normally distributed. I did because they always affect the quality.


# Bivariate Plots Section

We are going to group our sample based on the quality of it. We will have the mean volatile.acidity and median volatile.acidity also the number of occurence.

```{r echo=FALSE, Bivariate_Group_By_Quality}

# Group by quality, and calculate the mean of volatile.acidity & median of volatile.acidity. 
# Then calculate number of them.

rw.quality <- rw %>%
  group_by(quality) %>%
  summarise(mean_volatile.acidity = mean(volatile.acidity),
            median_volatile.acidity = median(volatile.acidity),
            n = n()) %>%
  ungroup() %>%
  arrange(quality)

# Print result
rw.quality
```

And here is a summary about the new variable rw.quality

```{r echo=FALSE, Summary_Of_rw.quality}

summary(rw.quality)

```

```{r echo=FALSE, Quality_Vs_Volatile.Acidity}

grid.arrange(ggplot(aes(x=factor(quality), y=volatile.acidity), data=rw) +
  geom_jitter(alpha=0.5) +
  geom_boxplot(alpha=0.5, color='gray') +
  stat_summary(fun.y='median',
               geom='point',
               color='blue',
               shape=10,
               size=10) +
    xlab("Factor of Quality") +
    ylab("Volatile Acidity"))

```


Our dataset contains some outlier, so we used median. We can see that as the quality increses, the volatile acidity decreases.

```{r echo=FALSE, Correlation_Table}

# Print a correlation tabel. 
round(cor(subset(rw, select = -c(ranking, category))), 3)

```

We can see that there is a highe correlation between alcohol and quality. 

```{r echo=FALSE, Correlations_Diagram}

red_wine_sample <- subset(rw, select = -c(total.sulfur.dioxide, residual.sugar, free.sulfur.dioxide, sulphates, fixed.acidity, category, ranking))

ggpairs(red_wine_sample,
        wrap=list(size=4),
        lower=list(continuous=wrap('points',shape = I('.'))),
        upper=list(combo=wrap('box', outlier.shape = I('.'))))+
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        axis.ticks = element_blank(), 
        panel.border = element_rect(linetype = "dashed", colour = "blue", fill = NA))

```

The above visualization shows a strong correlation between quality & alcohol. Also, it shows negtibe correlation between volatile.acidity & quality and cetric.acid & pH.

*Now, we will check relation between pH and alcohol.*

```{r echo=FALSE, Correlation_Diagram}

# Draw a geom point to the relation between pH and alcohol.
ggplot(data=rw, aes(x=pH, y=alcohol)) +
  geom_point() +
  geom_smooth()

```

*The above visualization shows low correlation, so we are going to check it with correlation test*

```{r echo=FALSE, Correlation_Test_phAndAlcohol}

# Test correlation between pH and alcohol
cor.test(rw$pH, rw$alcohol)

```

As expected, now much correlation between them.

```{r echo=FALSE, density_And_alcohol}

# Draw a geom point to the relation between density and alcohol.
ggplot(data=rw, aes(x=density, y=alcohol)) +
  geom_point(alpha=0.3, size=2) +
  geom_smooth(method='lm', se=FALSE)

```

*The above visualization shows some correlation, so we are going to check it with correlation test*

```{r echo=FALSE, Correlation_Test_densityAndAlcohol}

# Test correlation between density and alcohol
cor.test(rw$density, rw$alcohol)

```

We can see that there is a negtive correlation between [density, alcohol].

```{r echo=FALSE, chlorides_And_alcohol}

# Draw the relation between chlorides and alcohol.
ggplot(data=rw, aes(x=chlorides, y=alcohol))+
  geom_point(alpha=0.5, size=1)+
  geom_smooth(method='lm', se=FALSE)+
  xlim(0.05,quantile(rw$volatile.acidity, 0.5))+
  ylim(9,quantile(rw$alcohol, 1))+
  scale_x_log10()
```

*The above visualization shows outlaier, so we are going to check it with correlation test*

```{r echo=FALSE, Correlation_Test_chloridesAndAlcohol}

# Test correlation between chlorides and alcohol
cor.test(rw$chlorides, rw$alcohol)

```

We can see that there is no much correlation between [chlorides, alcohol].



# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

1. Density with alcohol: When alcohol increases, density decreases.
2. Residual sugar with alcohol: weak.
3. Acidic with pH: weak.



### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The relationship between alcohol & density, When alcohol increases, density decreases.

### What was the strongest relationship you found?

I was expecting many strong relationships, but the dataset shows nothing.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

# Plot about density vs alcohol and coloring rankings 
ggplot(data=rw, aes(x=density, y=alcohol, color=ranking)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE)

```


```{r echo=FALSE, Multivariate_Plots_pH&Alcohol}

# Plot about pH vs alcohol and coloring rankings 
ggplot(data=rw, aes(x=pH, y=alcohol ,color=ranking)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE)

```

The above two visualizations show that the excellent wine has a higer alcohol.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

I can clrearly see that alcohol and density affact the ranking of the wine.

### Were there any interesting or surprising interactions between features?

Ranking with other variables.

------

# Final Plots and Summary


### Plot One

```{r echo=FALSE, Plot_One}

ggplot(data=rw, aes(x=as.factor(quality), fill=ranking)) +
  geom_bar(aes(y=100*..count../sum(..count..))) +
  xlab("Wine Quality") +
  ylab('Percentage') +
  scale_fill_discrete(name="Alcohol Ranking") +
  ggtitle('Wine Quality')

```

### Description One

We can see that most of the sample quality are between [5, 6].

### Plot Two
```{r echo=FALSE, Plot_Two}

ggplot(data=rw, aes(x=pH, y=density, color=ranking)) +
  geom_point(position='jitter', alpha=0.5) +
  geom_smooth(method='lm', se=FALSE) +
  ylab('Density') + 
  xlab('pH') +
  ggtitle('Density and pH for all Ranking') 
  

```

### Description Two

We can see that as the density increses, the pH decreses. Which meaen there is a negtive correlation. 

### Plot Three

```{r echo=FALSE, Plot_Three}

ggplot(data=rw, aes(x=alcohol, y=density, color=ranking)) +
  geom_point(alpha=0.3, size=0.2) +
  geom_jitter(alpha=0.5) +
  geom_smooth(method='lm', se=FALSE) +
  xlab('Alcohol') + 
  ylab('Density') +
  ggtitle('Density VS Alcohol for Rankings')

```

### Description Three

A strong correlation, as alcohol increase, density decrease. 

------

# Reflection

We can see that alcohol & density are the best indicators for the wine quality. This will help all people who drinks red wine to take better decisions based on chemical details.